---
title: "Stock analysis of top food & beverage companies"
author: "Elmer Ore"
date: "Aug 17th, 2020"
output: 
  ioslides_presentation:
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 2019 Top Food companies {.contents}

According to [Food Engineering Magazine](https://www.foodengineeringmag.com/2019-top-100-food-beverage-companies>), these are the top 5 Food and Beverage companies in the world:

```{r echo = FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
library(kableExtra)

companies <- data.frame(Company = c("Nestle", "Pepsico, Inc", "Anheuser-Busch InBev", "JBS", "Tyson Foods"),
                        food_sales = c(80195, 64661, 54619, 46790, 40052))

table <- companies %>%
  mutate(sales =  paste0('$',format(food_sales, format='f', big.mark = ',', 
                                                digits = 0, nsmall=0, scientific = FALSE))) %>%
  select (-food_sales)
  

table %>%
  kable(col.names=c("Company", "Food Sales")) %>%
  kable_styling() %>%
   footnote(general = "Food sales are listed in millions of US dollars.")

```

## Adjusted Price series {.contents}

```{r echo = FALSE, message=FALSE, warning=FALSE, results='hide', cache=TRUE}
library(quantmod)
library(dplyr)
library(tidyr)

# Getting stock numbers from Yahoo Finance.
symbols <- c("NSRGY", "PEP", "BUD", "JBSAY", "TSN")
getSymbols(symbols, src = "yahoo", from = "2016-01-01", to = "2020-05-25", auto.assign = TRUE)

dfNSRGY <- as.data.frame(NSRGY) %>%
  rename (price = tail(names(.),1)) %>%
  mutate(date=time(NSRGY), index="Nestle") %>%
  select(index, price, date)

dfPEP <- as.data.frame(PEP) %>%
  rename (price = tail(names(.),1)) %>%
  mutate(date=time(PEP), index="Pepsico") %>%
  select(index, price, date)

dfBUD <- as.data.frame(BUD) %>%
  rename (price = tail(names(.),1)) %>%
  mutate(date=time(BUD), index="Anheuser-Busch Inbev") %>%
  select(index, price, date)

dfJBSAY <- as.data.frame(JBSAY) %>%
  rename (price = tail(names(.),1)) %>%
  mutate(date=time(JBSAY), index="JBS") %>%
  select(index, price, date)

dfTSN <- as.data.frame(TSN) %>%
  rename (price = tail(names(.),1)) %>%
  mutate(date=time(TSN), index="Tyson Foods") %>%
  select(index, price, date)

#combining the data from all 5 companies in one dataset.
data <- bind_rows(dfNSRGY, dfPEP, dfBUD, dfJBSAY, dfTSN)

```

```{r echo = FALSE, message=FALSE, warning=FALSE, out.width="100%", out.height="100%"}
library(dplyr)
library(tidyr)
library(plotly)

pal <- c("#cccccc", "#969696", "red", "#636363", "#252525")
pal <- setNames(pal, c("Nestle", "Pepsico", "Anheuser-Busch Inbev", "JBS", "Tyson Foods"))

fig1 <- plot_ly(data, type="scatter", mode="markers",
        x=~date, y=~price, color=~index,
        colors =  pal, 
        marker = list(size = 4))#,
       # width = 800,
        #height = 500)

fig1 <- fig1 %>%  
  layout(title= "Adjusted stock price of top food and beverage brands",
         font = list(
           family = "Montserrat, sans-serif",
           size = 20),
         yaxis = list(zeroline = FALSE, 
                      tickprefix="Price ", 
                      showtickprefix = 'last',
                      title = "",
                      font = list(
                        family = "Montserrat, sans-serif",
                        size = 14)),
         xaxis = list(zeroline = FALSE),
         showlegend = FALSE,
         margin = 1
         )

lblNestle <- list(
  xref = 'paper',
  x = 0.95,
  y = dfNSRGY[nrow(dfNSRGY),2],  #price
  xanchor = 'left',
  yanchor = 'middle',
  text = 'Nestle',
  font = list(color = "#cccccc"), #'rgba(67,67,67,1)'),
  showarrow = FALSE)

lblPepsico <- lblNestle
lblPepsico$y <- dfPEP[nrow(dfPEP),2]
lblPepsico$text <- "Pepsico"
lblPepsico$font$color <- "#969696"

lblAnheuser <- lblPepsico
lblAnheuser$y <- dfBUD[nrow(dfBUD),2]
lblAnheuser$text <- "Anheuser-Busch Inbev"
lblAnheuser$font$color <- "red"

lblJBS <- lblPepsico
lblJBS$y <- dfJBSAY[nrow(dfJBSAY),2]
lblJBS$text <- "JBS"
lblJBS$font$color <- "#636363"

lblTyson <- lblPepsico
lblTyson$y <- dfTSN[nrow(dfTSN),2]
lblTyson$text <- "Tyson Foods"
lblTyson$font$color <- "#252525"

fig1 <- fig1 %>% 
  layout(annotations = lblNestle) %>%
  layout(annotations = lblPepsico) %>%
  layout(annotations = lblAnheuser) %>%
  layout(annotations = lblJBS) %>%
  layout(annotations = lblTyson)
  
fig1
```

## Trends from Anheuser-Busch Inbev prices {.contents}

```{r echo = FALSE, message=FALSE, warning=FALSE, out.width="100%", out.height="100%"}
library(dplyr)
library(zoo)
library(plotly)

# calculating moving averages
dfBUD_mm <- filter(dfBUD, date >= "2019-01-01") %>% select (-index)

dfBUD_mm <- dfBUD_mm %>%
  mutate(mm10 = rollmean(dfBUD_mm$price, 10, fill = list(NA, NULL, NA), align = "right")) %>%
  mutate(mm30 = rollmean(dfBUD_mm$price, 30, fill = list(NA, NULL, NA), align = "right"))

pal2 <- c("#969696", "firebrick4", "darkcyan")
pal2 <- setNames(pal2, c("price", "mm10", "mm30"))

fig2 <- plot_ly(dfBUD_mm, type="scatter", mode="lines",
                x=~date, y=~price,
                name="price",
                colors = pal2, 
                color= "price",
                height = 500)

fig2 <- fig2 %>% 
  add_trace(y = ~mm10, name = '10 days', color="mm10") %>%
  add_trace(y = ~mm30, name = '30 days', color="mm30")

fig2 <- fig2 %>%  
  layout(title= "Moving averages of Anheuser-Busch's adjusted price",
         font = list(
           family = "Montserrat, sans-serif",
           size = 20), 
         yaxis = list(zeroline = FALSE, 
                      tickprefix="Price ", 
                      showtickprefix = 'last',
                      title = "",
                      font = list(
                        family = "Montserrat, sans-serif",
                        size = 14)),
         xaxis = list(zeroline = FALSE),
         margin = 10
  )

fig2
```


## Returns from Anheuser-Busch Inbev {.contents}

```{r echo = FALSE, message=FALSE, warning=FALSE, out.width="100%", out.height="100%"}
library(dplyr)
library(plotly)

# calculating returns
dfBUD_ret <- as.data.frame(diff(log(dfBUD$price)) )
dfBUD_ret <- mutate(dfBUD_ret, returns=diff(log(dfBUD$price))) %>% select(returns)
dfBUD_ret <- bind_rows(data.frame(returns=NA), dfBUD_ret)
dfBUD_ret <- bind_cols(dfBUD, returns = dfBUD_ret)

fig3 <- plot_ly(dfBUD_ret, type="scatter", mode="lines",
                x=~date, y=~returns,
                height = 500)

fig3 <- fig3 %>%  
  layout(title= "Anheuser-Busch's return series",
         font = list(
           family = "Montserrat, sans-serif",
           size = 20), 
         yaxis = list(zeroline = FALSE, 
                      tickprefix="Returns ", 
                      showtickprefix = 'last',
                      title = "",
                      font = list(
                        family = "Montserrat, sans-serif",
                        size = 14)),
         xaxis = list(zeroline = FALSE),
         margin = 10
  )

fig3
```


## Summary {.contents}

- Anheuser-Busch InBev adjusted stock prices have the worst decreased in value between 2016 until May 2020 between the top 5 companies in the ranking.
- Using smooth series we can see that there are 8 times where the series crossed themselves and 2 overlaps. These are moments where an evaluation of buy or sell stocks should be considered.
- The returns of Anheuser-Busch InBev shows increased volatility since March 2020, when the coronavirus epidemic became global.

## References {.contents}

- [Food Engineering Magazine](https://www.foodengineeringmag.com/2019-top-100-food-beverage-companies>) 2019 Top 100 Food & Beverage Companies.
- [Introduction to stock analysys in R](https://lamfo-unb.github.io/2017/07/22/intro-stock-analysis-1/) by Gustavo Monteiro
- [R Figure reference](https://plotly.com/r/reference/) from Plotly.
- [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/) by Yihui Xie, J. J. Allaire, Garrett Grolemund. 
